---
title: "TRD DIABLO: Multiband with 3 networks and Maxdiff"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2025-05-23"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/data/leey19/sandbox/TRD_DIABLO/DIABLO_conn/LCMV/kong2022_200_2/subtype"))

```



## max difference model

### Setup the environment

The input sample size was 178, 86 for HCs and 92 for MDDs.

```{r, include=FALSE}

rm(list = ls())

#BiocManager::install('mixOmics')
library(mixOmics) # import the mixOmics library
library(dplyr)
library(moments)
library(robustHD)

set.seed(4982)

#file.choose()
setwd("/data/MoodMEG/trd_project/sample_JSloan/")
# setwd("/Users/leey19/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/Documents/NIH_Postdoc/Sandbox/TRD_DIABLO/results/LCMV/kong2022_200_2/")

ids <- read.csv("TRD_DIABLO_group_MDDtype.csv")

table(ids$Group)

```

### Setup the values

#### Setup the X1 value using the Default Mode Network (DMN) correlatioin coefficient

Potentially outlier: CWJYSXBI

```{r, warning=FALSE, message=FALSE}

X0 <- read.csv("envcorr_kong200_all_default_FIN.csv",header=T)

transform_group <- function(data) {
  numeric_data <- data[, sapply(data, is.numeric)]
  log_data <- log(numeric_data)
  robust_scale <- function(x) {
    med <- median(x)
    iqr <- IQR(x)
    scaled_data <- (x - med) / iqr
    scaled_data[scaled_data > 3] <- 3
    scaled_data[scaled_data < -3] <- -3
    return(scaled_data)
  }
  scaled_data <- apply(log_data, 2, robust_scale)
  return(scaled_data)
}

X0.1 <- X0 %>% select(where(~ !any(is.na(.))))
X0.1 <- X0.1[order(X0.1$ID),]
X0.1 <- merge(X0.1, ids, by="ID")
X0.2_list <- split(X0.1, X0.1$Group)
X0.2_transformed <- lapply(X0.2_list, function(group_data) {
    group_data_sub <- group_data[, !(names(group_data) %in% c("ID", "Group"))]
    transformed_data <- transform_group(group_data_sub)
    data.frame(ID = group_data$ID, transformed_data)
})
X0.3 <- do.call(rbind, X0.2_transformed)
rownames(X0.3) <- X0.3$ID

X1 <- X0.3


```

#### Setup the X2 value using the Executive Control Network (ECN) correlatioin coefficient

Potentially outlier: DIIOFUND

```{r, warning=FALSE, message=FALSE, include=FALSE}


X0 <- read.csv("envcorr_kong200_all_control_FIN.csv",header=T)

X0.1 <- X0 %>% select(where(~ !any(is.na(.))))
X0.1 <- X0.1[order(X0.1$ID),]
X0.1 <- merge(X0.1, ids, by="ID")
X0.2_list <- split(X0.1, X0.1$Group)
X0.2_transformed <- lapply(X0.2_list, function(group_data) {
  group_data_sub <- group_data[, !(names(group_data) %in% c("ID", "Group"))]
  transformed_data <- transform_group(group_data_sub)
  data.frame(ID = group_data$ID, transformed_data)
})
X0.3 <- do.call(rbind, X0.2_transformed)
rownames(X0.3) <- X0.3$ID

X2 <- X0.3


```



#### Setup the X3 value using the Sailience Network (SN) correlatioin coefficient

Potentially outlier: DIIOFUND

```{r, warning=FALSE, message=FALSE, include=FALSE}

X0 <- read.csv("envcorr_kong200_all_salience_FIN.csv",header=T)

X0.1 <- X0 %>% select(where(~ !any(is.na(.))))
X0.1 <- X0.1[order(X0.1$ID),]
X0.1 <- merge(X0.1, ids, by="ID")
X0.2_list <- split(X0.1, X0.1$Group)
X0.2_transformed <- lapply(X0.2_list, function(group_data) {
  group_data_sub <- group_data[, !(names(group_data) %in% c("ID", "Group"))]
  transformed_data <- transform_group(group_data_sub)
  data.frame(ID = group_data$ID, transformed_data)
})
X0.3 <- do.call(rbind, X0.2_transformed)
rownames(X0.3) <- X0.3$ID

X3 <- X0.3



```


#### Setup the X4 value using the Doral Attentional Network (DAN) correlatioin coefficient

Potentially outlier: GMZHYQVF

```{r, warning=FALSE, message=FALSE, include=FALSE}

X0 <- read.csv("envcorr_kong200_all_dorsatt_FIN.csv",header=T)

X0.1 <- X0 %>% select(where(~ !any(is.na(.))))
X0.1 <- X0.1[order(X0.1$ID),]
X0.1 <- merge(X0.1, ids, by="ID")
X0.2_list <- split(X0.1, X0.1$Group)
X0.2_transformed <- lapply(X0.2_list, function(group_data) {
  group_data_sub <- group_data[, !(names(group_data) %in% c("ID", "Group"))]
  transformed_data <- transform_group(group_data_sub)
  data.frame(ID = group_data$ID, transformed_data)
})
X0.3 <- do.call(rbind, X0.2_transformed)
rownames(X0.3) <- X0.3$ID

X4 <- X0.3


```



#### Setup the X5 value using the Motor Network (MN) correlatioin coefficient

Potentially outlier: FQTIDXBD & WNKOWTSP

```{r, warning=FALSE, message=FALSE, include=FALSE}

X0 <- read.csv("envcorr_kong200_all_motor_FIN.csv",header=T)

X0.1 <- X0 %>% select(where(~ !any(is.na(.))))
X0.1 <- X0.1[order(X0.1$ID),]
X0.1 <- merge(X0.1, ids, by="ID")
X0.2_list <- split(X0.1, X0.1$Group)
X0.2_transformed <- lapply(X0.2_list, function(group_data) {
  group_data_sub <- group_data[, !(names(group_data) %in% c("ID", "Group"))]
  transformed_data <- transform_group(group_data_sub)
  data.frame(ID = group_data$ID, transformed_data)
})
X0.3 <- do.call(rbind, X0.2_transformed)
rownames(X0.3) <- X0.3$ID

X5 <- X0.3


```


#### Setup the X6 value using the Visual Network (VN) correlatioin coefficient

Potentially outlier: ETWNABKX

```{r, warning=FALSE, message=FALSE, include=FALSE}

X0 <- read.csv("envcorr_kong200_all_visual_FIN.csv",header=T)

X0.1 <- X0 %>% select(where(~ !any(is.na(.))))
X0.1 <- X0.1[order(X0.1$ID),]
X0.1 <- merge(X0.1, ids, by="ID")
X0.2_list <- split(X0.1, X0.1$Group)
X0.2_transformed <- lapply(X0.2_list, function(group_data) {
  group_data_sub <- group_data[, !(names(group_data) %in% c("ID", "Group"))]
  transformed_data <- transform_group(group_data_sub)
  data.frame(ID = group_data$ID, transformed_data)
})
X0.3 <- do.call(rbind, X0.2_transformed)
rownames(X0.3) <- X0.3$ID

X6 <- X0.3

```


#### Setup the X and Y values using the clinician-driven diagnosis

from 178

We merged the X1 to X6 while log-scaling values. The dimension of the available data are:

1. DMN: 178 individuals with 253 variables (unique correlation coefficients)
2. ECN: 178 individuals with 120 variables (unique correlation coefficients)
3. SN: 178 individuals with 105 variables (unique correlation coefficients)
4. DAN: 178 individuals with 66 variables (unique correlation coefficients)
5. MN: 178 individuals with 66 variables (unique correlation coefficients)
6. VN: 178 individuals with 105 variables (unique correlation coefficients)
8. Y: RISC groups ( for 86 HC; for 92 MDD)



```{r, warning=FALSE, message=FALSE, include=FALSE}


# Y_name <- X1[,1]
# 
# rownames(X1) <- Y_name
# rownames(X2) <- Y_name
# rownames(X3) <- Y_name
# rownames(X4) <- Y_name
# rownames(X5) <- Y_name
# rownames(X6) <- Y_name
# 
# library(ExPosition)
# res <- epPCA(X6[-1], DESIGN=Y_name)
# sort(res$ExPosition.Data$eigs)
# sort(res$ExPosition.Data$t)
# sort(res$ExPosition.Data$fi[,1])

#X1: IFU/LNK/QZU/OEN/ITV/JBG/BDP
#X2: IFU/LNK/QZU/OEN/XBG/ITV/TUJ
#X3: IFU/LNK/QZU/ITV/OEN
#X4: IFU/LNK/ITV/XBG/QZU/OEN
#X5: IFU/LNK/QZU/ITV/OEN
#X6: IFU/LNK/XBG/CTX/OEN/JBG/ITV

# outlier: IFU/LNK/ITV/XBG/QZU/OEN


# ### Omit the participants
# # same individual drop: WNKOWTSP




X1.1 <- X1 %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')
X2.1 <- X2 %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')
X3.1 <- X3 %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')
X4.1 <- X4 %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')
X5.1 <- X5 %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')
X6.1 <- X6 %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')
Y.1 <- ids %>%
  filter(ID != 'WNKOWTSP' & ID != 'IFUOWLYC' & ID != 'QZUBAVVA' & ID != 'DWYASHSC' & ID != 'FQTIDXBD'& ID != 'LNKRYMPE'& ID != 'OENJHTBX'& ID != 'IBTXSVUE'& ID != 'TUJTIRFK'& ID != 'ITVESALZ')




X1.2 <- X1.1[! names(X1.1) %in% c('ID')]
X2.2 <- X2.1[! names(X2.1) %in% c('ID')]
X3.2 <- X3.1[! names(X3.1) %in% c('ID')]
X4.2 <- X4.1[! names(X4.1) %in% c('ID')]
X5.2 <- X5.1[! names(X5.1) %in% c('ID')]
X6.2 <- X6.1[! names(X6.1) %in% c('ID')]



#######

X <- list(DMN = X1.2, ECN = X2.2, 
          SN = X3.2)

Y <- Y.1$Group

```


#### Setup design matrix

3 networks

```{r}


design = matrix(0.1, ncol = length(X), nrow = length(X), 
                dimnames = list(names(X), names(X)))


diag(design) = 0 # set diagonal to 0s


```

#### Tuning the model

```{r}
basic.diablo.model = block.splsda(X = X, Y = Y, ncomp = 5, design = design) 

perf.diablo = perf(basic.diablo.model, validation = 'Mfold', 
                   folds = 10, nrepeat = 10) 

plot(perf.diablo) # plot output of tuning

# set the optimal ncomp value
ncomp = perf.diablo$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"]

# show the optimal choice for ncomp for each dist metric
perf.diablo$choice.ncomp$WeightedVote

```



#### Setup and run the DIABLO model

We conducted an analysis using the DIABLO model, with 10 folds and LOOCV approach.



```{r}


#note: We choose the optimal number of variables to select in each data set
#using the tune.block.splsda() function

# set grid of values for each component to test
# DMN/ECN/SN/...
test.keepX = list(DMN =c(1:9, seq(10, 30, 2), 35),
                  ECN =c(1:9, seq(10, 18, 2), seq(20,25,5),seq(26,34,2)),
                  SN =c(1:9, seq(10, 18, 2), seq(20,25,5),seq(26,34,2)))


 
# library(BiocParallel)
# BPPARAM <- BiocParallel::MulticoreParam(workers = parallel::detectCores()-1)
# print(BPPARAM)
# 
# tune.RISC = tune.block.splsda(X = X, Y = Y,
#                               test.keepX = test.keepX, design = design,
#                               validation = 'Mfold', folds=5,  nrepeat=100,
#                               dist = "centroids.dist", ncomp = 2,
#                               seed='4982',BPPARAM=BPPARAM)
# 
# 
# tune.RISC$choice.keepX


# saveRDS(tune.RISC, file = "tuneTRD_DIABLO_kong200_3network_MaxDiff.Rds")

```



### Results

#### Check the optimal values of features to retain in each X value

Note: I read the outcome here.


```{r, message=FALSE}

tune.RISC <- readRDS(file = "tuneTRD_DIABLO_kong200_3network_MaxDiff.Rds")

list.keepX = tune.RISC$choice.keepX # set the optimal values of features to retain
list.keepX

result.sparse.diablo.risc <-  block.splsda(X, Y, keepX = list.keepX,
                                           design = design) 

# the features selected to form the first component
### 2 groups (Group HC/DEP; tuneTRD_DIABLO_PLS.Rds) with a PLS approach
#DMN: 1
#ECN: 5
#SN: 1

```

#### Check the selected features of DMN (component=1)

Here, we can see the list of variables (in component 1 of DMN).

```{r}
selectVar(result.sparse.diablo.risc, block = 'DMN', comp = 1)$DMN$name 

```

#### Check the selected features of DMN (component=2)

Here, we can see the list of variables (in component 2 of DMN).

```{r}
selectVar(result.sparse.diablo.risc, block = 'DMN', comp = 2)$DMN$name 

```

#### Check the selected features of ECN (dimension=1)

Desc

```{r}
selectVar(result.sparse.diablo.risc, block = 'ECN', comp = 1)$ECN$name 

```

#### Check the selected features of ECN (dimension=2)

Desc

```{r}
selectVar(result.sparse.diablo.risc, block = 'ECN', comp = 2)$ECN$name 

```

#### Check the selected features of SN (dimension=1)

Desc

```{r}
selectVar(result.sparse.diablo.risc, block = 'SN', comp = 1)$SN$name 

```

#### Check the selected features of SN (dimension=2)

Desc

```{r}
selectVar(result.sparse.diablo.risc, block = 'SN', comp = 2)$SN$name 

```

#### Check the model performance using the package

The model performance was estimated using the "RVAideMemoire" in R.

* 2 groups (Group TRD/HC; tuneTRD_DIABLO_3network_MaxDiff.Rds)
* 999 permutations
* CER = .584
* p-value = .938

```{r}
#library(RVAideMemoire)
#DIABLO.test(result.sparse.diablo.risc)



```

#### Check the model performance using the AUC

dsalkdsajdklsaj

```{r}
# run repeated CV performance evaluation
perf.diablo = perf(result.sparse.diablo.risc, validation = 'Mfold', folds = 5,
                   nrepeat = 100, auc = T,
                   dist = 'centroids.dist') 

perf.diablo$MajorityVote.error.rate

perf.diablo$WeightedVote.error.rate

perf.diablo$auc

```

###### DMN (component=1)

```{r}
auc.splsda = auroc(result.sparse.diablo.risc, roc.block = "DMN", 
                   roc.comp = 1, print = FALSE)

auc.splsda

```

###### DMN (component=2)

```{r}
auc.splsda = auroc(result.sparse.diablo.risc, roc.block = "DMN", 
                   roc.comp = 2, print = FALSE)

auc.splsda

```

###### ECN (component=1)

```{r}
auc.splsda = auroc(result.sparse.diablo.risc, roc.block = "ECN", 
                   roc.comp = 1, print = FALSE)

auc.splsda

```

###### ECN (component=2)

```{r}
auc.splsda = auroc(result.sparse.diablo.risc, roc.block = "ECN", 
                   roc.comp = 2, print = FALSE)

auc.splsda

```

###### SN (component=1)

```{r}
auc.splsda = auroc(result.sparse.diablo.risc, roc.block = "SN", 
                   roc.comp = 1, print = FALSE)

auc.splsda

```

###### SN (component=2)

```{r}
auc.splsda = auroc(result.sparse.diablo.risc, roc.block = "SN", 
                   roc.comp = 2, print = FALSE)

auc.splsda

```

### Plots

#### Correlation among Xs (component=1)

dsalkdsajdklsaj

```{r, out.width = '100%'}
plotDiablo(result.sparse.diablo.risc, ncomp = 1)

```

#### Correlation among Xs (component=2)

dsalkdsajdklsaj

```{r, out.width = '100%'}
plotDiablo(result.sparse.diablo.risc, ncomp = 2)

```

#### Individual comparions between Xs
dsalkdsajdklsaj

```{r, out.width = '100%'}
plotIndiv(result.sparse.diablo.risc, ind.names = FALSE, legend = TRUE, 
          title = 'DIABLO Sample Plots')

```

#### Arrow plot

dsalkdsajdklsaj

```{r}
plotArrow(result.sparse.diablo.risc, ind.names = F, legend = TRUE, 
          title = 'DIABLO')
```


#### Correlation Circle Plot 1

dsdskjdsk

```{r, out.width = '100%'}
plotVar(result.sparse.diablo.risc, var.names = FALSE, 
        style = 'graphics', legend = TRUE,
        pch = c(16, 17, 15), cex = c(2,2,2), 
        col = c('darkorchid', 'brown1', 'lightgreen'))

```

#### Correlation Circle Plot 2

dsdskjdsk

```{r, out.width = '100%'}
circosPlot(result.sparse.diablo.risc, cutoff = 0.6, line = TRUE,
           color.blocks= c('darkorchid', 'brown1', 'lightgreen'),
           color.cor = c("chocolate3","grey20"), size.labels = 1.5)
```


#### plot loadings of component 1

dsdskjdsk

```{r, echo=FALSE, fig.cap="Component 1", out.width = '100%'}

plotLoadings(result.sparse.diablo.risc, comp = 1, contrib = 'max', method = 'median',size.name = .6)


```




#### plot loadings of component 2

dsdskjdsk

```{r, echo=FALSE, fig.cap="Component 2", out.width = '100%'}

plotLoadings(result.sparse.diablo.risc, comp = 2, contrib = 'max', method = 'median',size.name = .6)


```


#### save contingency output

dsdskjdsk

```{r, echo=T}

library(dplyr)
library(pROC)

res1 <- as.data.frame(perf.diablo$AveragedPredict.class$nrep10)
res1$id <- rownames(res1)

res2 <- as.data.frame(result.sparse.diablo.risc$Y)
names(res2)[1] <- "Y"

res3 <- cbind(res1, res2)

res3 <- res3 %>%
  mutate(Y.num  = ifelse(Y=="HC",ifelse(Y=="MDD",NA,1),0),
         dim1.num  = ifelse(dim1=="HC",ifelse(dim1=="MDD",NA,1),0),
         dim2.num  = ifelse(dim2=="HC",ifelse(dim2=="MDD",NA,1),0)) %>%
  select(id, Y.num, dim1.num, dim2.num)

write.csv(res3, "TRD_DIABLO_cont_3network_maxdiff_anx_broad.csv", row.names = F)


```







